### 2020-07-14 latex/fig/dc.tex
DC Mall Figure

\begin{tikzpicture}
    \node[inner sep=0pt] (original) at (0,0)
        {\includegraphics[height=0.20\textwidth]{data/DC/0.png}};
        \node[inner sep=0pt] (1) at (-3,-2)
            {\includegraphics[height=0.20\textwidth]{data/DC/1.png}};
            \node[inner sep=0pt] (3) at (-1.6,-4)
                {\includegraphics[height=0.20\textwidth]{data/DC/3.png}};
                \node[inner sep=0pt] (7) at (-2.5,-6)
                    {\includegraphics[height=0.20\textwidth]{data/DC/7.png}};
                \node[inner sep=0pt] (8) at (-0.7,-6)
                    {\includegraphics[height=0.20\textwidth]{data/DC/8.png}};
            \node[inner sep=0pt] (4) at (-4.4,-4)
                {\includegraphics[height=0.20\textwidth]{data/DC/4.png}};
        \node[inner sep=0pt] (2) at (3,-2)
            {\includegraphics[height=0.20\textwidth]{data/DC/2.png}};
            \node[inner sep=0pt] (5) at (4.4,-4)
                {\includegraphics[height=0.20\textwidth]{data/DC/3.png}};
            \node[inner sep=0pt] (6) at (1.6,-4)
                {\includegraphics[height=0.20\textwidth]{data/DC/4.png}};
                \node[inner sep=0pt] (13) at (0.7,-6)
                    {\includegraphics[height=0.20\textwidth]{data/DC/3.png}};
                \node[inner sep=0pt] (14) at (2.5,-6)
                    {\includegraphics[height=0.20\textwidth]{data/DC/4.png}};
\end{tikzpicture}