### 2020-07-15 latex/plots.tex
realworld hierarchical and level breakdown plots

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_rank2_strongscaling.pdf}
\caption{Relative Time Breakdown for Rank-2 NMF \cref{alg:parrank2nmf} Strong Scaling at Level-0}
\label{fig:synrank2strongscaling}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_rank2_speedup.pdf}
\caption{Strong Scaling Speedup for Rank-2 NMF \cref{alg:parrank2nmf} at Level-0}
\label{fig:synrank2speedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_hierarchical_speedup.pdf}
\caption{Strong Scaling Speedup for Hierarchical Clustering. The total time taken for complete Hierarchical Clustering \cref{alg:hiernmf2}}
\label{fig:synhierspeedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_hier_strongscaling.pdf}
\caption{Relative Time Breakdown for Hierarchical Clustering. The plot uses the total time taken for complete Hierarchical Clustering \cref{alg:hiernmf2}}
\label{fig:synhierstrongscaling}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_sequential_level_breakdown.pdf}
\caption{Sequential Breakdown plot for levels 0 to 5. y-axis is absolute wall clock time in seconds}
\label{fig:seqlevelbreakdown}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_parallel_level_breakdown.pdf}
\caption{Parallel Breakdown plot for levels 0 to 5 on 40 nodes. y-axis is absolute wall clock time in seconds}
\label{fig:parallellevelbreakdown}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/synthetic_rank2_weakscaling.pdf}
\caption{Relative Time Breakdown for Rank-2 NMF \cref{alg:parrank2nmf} Weak Scaling at Level-0}
\label{fig:synrank2weakscaling}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_rank2_strongscaling.pdf}
\caption{Relative Time Breakdown for Rank-2 NMF \cref{alg:parrank2nmf} Strong Scaling at Level-0 on Realworld Data}
\label{fig:rwrank2strongscaling}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_rank2_speedup.pdf}
\caption{Strong Scaling Speedup for Rank-2 NMF \cref{alg:parrank2nmf} at Level-0 on Realworld Data}
\label{fig:rwrank2speedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_hierarchical_speedup.pdf}
\caption{Strong Scaling Speedup for Hierarchical Clustering on Realworld Data. The total time taken for complete Hierarchical Clustering \cref{alg:hiernmf2}}
\label{fig:rwhierspeedup}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_hier_strongscaling.pdf}
\caption{Relative Time Breakdown for Hierarchical Clustering on Realworld. The plot uses the total time taken for complete Hierarchical Clustering \cref{alg:hiernmf2}}
\label{fig:rwhierstrongscaling}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_sequential_level_breakdown.pdf}
\caption{Parallel Breakdown plot for levels 0 to 4 on 10 nodes. y-axis is absolute wall clock time in seconds}
\label{fig:rwseqlevelbreakdown}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=2in, width=\columnwidth]{plots/realworld_parallel_level_breakdown.pdf}
\caption{Parallel Breakdown plot for levels 0 to 5 on 80 nodes. y-axis is absolute wall clock time in seconds}
\label{fig:rwparallellevelbreakdown}
\end{center}
\end{figure}