### 2020-07-16 latex/tex/conclusion.tex

\section{Conclusion}

\begin{itemize}
	\item consider sparse case
	\item consider parallelizing across tree
	\item consider redistribution for 2D
\end{itemize}