### 2020-07-16 latex/tex/conclusion.tex
moved comment from intro to conclusion

% !TEX root = ../paper.tex

\section{Conclusion}

\begin{itemize}
	\item consider sparse case
	\item consider parallelizing across tree
	\item consider redistribution for 2D
	\item compare performance with flat NMF and also employ DC-NMF
\end{itemize}