### 2020-11-13 latex/paper.tex
adding description of rank scaling experiment, still need to add figure

\documentclass[10pt, conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\usepackage{cite,doi}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm,algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{cleveref}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{ourmacros}
\usepackage{subcaption}

\usepackage[subtle,tracking=normal,leading=tight,leadingfraction=0.98, indent=tight, wordspacing=normal, bibliography=tight]{savetrees}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage{paralist}
\renewenvironment{itemize}[1]{\begin{compactitem}#1}{\end{compactitem}}

\newcommand{\LM}[1]{\textcolor{blue}{\textbf{Lawton}: #1}}    
\newcommand{\GB}[1]{\textcolor{red}{\textbf{Grey}: #1}}
\newcommand{\RK}[1]{\textcolor{blue}{\textbf{Ramki}: #1}}

\newcommand{\hyper}{DC-HYDICE}
\newcommand{\image}{SIIM-ISIC}

\begin{document}

\title{Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization}

\author{\IEEEauthorblockN{Lawton Manning and Grey Ballard}
\IEEEauthorblockA{Wake Forest University \\
Winston-Salem, NC, USA \\
\{mannlg15,ballard\}@wfu.edu}
\and
\IEEEauthorblockN{Ramakrishnan Kannan}
\IEEEauthorblockA{Oak Ridge National Laboratory \\
Oak Ridge, TN, USA \\
kannanr@ornl.gov}
\and
\IEEEauthorblockN{Haesun Park}
\IEEEauthorblockA{Georgia Institute of Technology \\
Atlanta, GA, USA \\
hpark@cc.gatech.edu}
}

\maketitle

\input{tex/abstract.tex}

\begin{IEEEkeywords}
low-rank approximation, distributed-memory parallel algorithms, scalable clustering
\end{IEEEkeywords}

\input{tex/introduction.tex}

\input{tex/prelims.tex}

\input{tex/algorithm.tex}

\input{tex/analysis.tex}

\input{tex/experiments.tex}

\input{tex/conclusion.tex}

%Add this back in for camera ready
\input{tex/ack.tex}

\bibliographystyle{plainurl}
\bibliography{paper}

\end{document}